### 2011-12-11 plos_template.tex
ups, added statsmonkey
wrote draft on methods of survey and better draft on software-methods

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{openSNP - Crowdsourcing Genome Wide Association Studies}
}
% Insert Author names, affiliations and corresponding author email.
\\
Bastian Greshake$^{1,\ast}$, 
Philipp Bayer$^{2}$, 
Fabian Zimmer$^{3,}$
\\
\bf{1} Bastian Greshake Frankfurt am Main, Germany
\\
\bf{2} Philipp Bayer somewhere, Australia
\\
\bf{3} Fabian Zimmer M\"unster, Germany
\\
$\ast$ E-mail: info@opensnp.org
\end{flushleft}

@@

Although companies like 23andMe are willing to contribute to science it is not easy for individual scientists to get hold of the data. This arises mainly due to privacy concerns of the customers. Nevertheless there are individual customers who are willingly sharing their data. Most do though by uploading it to their personal website or to software repositories like GitHub. While this is makes it possible to use the data, it requires a lot of work to keep track of all new genotyping data that is available to the public. While projects like the SNPedia try to keep track of all the files, this still does not allow to perform GWAS, as the phenotypic information is not attached to the genetic information. Projects that attach the phenotype to the genetic information, like the Personal Genome Project, still don't allow for an easy re-use of the data.

@@

Those participants who are not planning to get genotyped are more likely to agree to not share their data, because they fear discrimination (mean difference = 1.52, SE = 0.322, p < 0.001), because they feel it is a breach of their privacy (mean difference = 1.871, SE = 0.324, p < 0.001), because they fear consequences for their family (mean difference = 1.146, SE = 0.32, p = 0.001) and because they fear personalized advertising (mean difference =  1.112, SE = 0.357, p = 0.006). 
\subsection*{Subsection 2}

@@

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}

>>>>

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{openSNP - Crowdsourcing Genome Wide Association Studies}
}
% Insert Author names, affiliations and corresponding author email.
\\
Bastian Greshake$^{1,\ast}$, 
Philipp Bayer$^{2}$, 
Fabian Zimmer$^{3}$,
Julia Reda$^{4}$
\\
\bf{1} Bastian Greshake Frankfurt am Main, Germany
\\
\bf{2} Philipp Bayer somewhere, Australia
\\
\bf{3} Fabian Zimmer M\"unster, Germany
\\
\bf{4} Julia Reda, Mainz, Germany
\\
$\ast$ E-mail: info@opensnp.org
\end{flushleft}

@@

Although companies like 23andMe are willing to contribute to science it is not easy for individual scientists to get hold of the data. This arises mainly due to privacy concerns of the customers. Nevertheless there are individual customers who are willingly sharing their data. Most do though by uploading it to their personal website or to software repositories like GitHub. While this is makes it possible to use the data, it requires a lot of work to keep track of all new genotyping data that is available to the public. While projects like the SNPedia (10.1093/nar/gkr798) try to keep track of all the files, this still does not allow to perform GWAS, as the phenotypic information is not attached to the genetic information. Projects that attach the phenotype to the genetic information, like the Personal Genome Project, still don't allow for an easy re-use of the data.

@@

Those participants who are not planning to get genotyped are more likely to agree to not share their data, because they fear discrimination (mean difference = 1.52, SE = 0.322, p < 0.001), because they feel it is a breach of their privacy (mean difference = 1.871, SE = 0.324, p < 0.001), because they fear consequences for their family (mean difference = 1.146, SE = 0.32, p = 0.001) and because they fear personalized advertising (mean difference =  1.112, SE = 0.357, p = 0.006). 
\subsection*{openSNP platform}

@@

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}
\subsection*{Survey on Sharing Genetic Information}
Done in Google Docs, questions on age, chromosomal sex, ethnicity, if they are already customer of a DTC company/planning to become one/don't plan to become one. If they are already a customer: Are they already sharing and are they already sharing phenotypic information? How strong on a scale of 1 = strongly disagree to  5 = strongly agree how do they feel about reasons to share their genetic data because... a) they want to help scientists b) get a personal benefit out of it (new treatments for illnesses) c) because of curiosity d) because it might bring relevant, personalized advertising. How strong on a scale of 1 = strongly disagree to 5 = strongly agree would they feel about not sharing their data because... a) they fear discrimination b) it is a breach of their privacy c) because of negative side-effects for their next of kin d) because they fear personalized spam.

The results were analyzed in somestatisticstool.

\subsection*{Technical realization of the platform}
The main platform is realized using the web-framework Ruby on Rails. Postgres is used as the main database backend for Rails. The database stores genotyping results, phenotypic information of the users, literature results of Mendeley, the Public Library of Science and summaries on SNPs which can be found in SNPedia.

The literature database of Mendeley is queried using the REST API, which delivers results in JSON. The literature database of the Public Library of Science is queried using the respective REST API, which delivers results in XML. Summaries on SNPs are provided by SNPedia, through querying the content through the MediaWiki-API.

Processes with a longer runtime, such as parsing the genotyping results, creating archives of results which are be mailed to users and queries to external resources are handled using the ruby-gem Resque and a Redis-server. Search-features on the platform itself are implemented using SOLR and the ruby-gem sunspot. Additionally data can be requested from openSNP using the Distributed Annotation System. The data for this is stored in a mySQL-database, the delivery of the data is done by using ProServer, which was modified by Gel et al. for use in easyDAS.

The source code of openSNP is published under Creative Commons BY-SA 3.0 and can be downloaded at http://github.com/gedankenstuecke/snpr. The genetical and phenotypical data is licensed under Creative Commons Zero.
